---
title: "4031Project_MohanOHareWallace"
author: "Ved Mohan"
date: "11/21/2020"
output:
  word_document: default
  html_document: default
---
# Introduction
Gentrification is defined as the transformation of areas inhabited by historically marginalized groups into those used by the dominant groups. Historically, this has served as a reflection of class and ethnic marginalization. The involuntary household movement caused by factors such as affordability and rising rent costs is essential to track in rapidly developing urban centers such as Atlanta.

This project will detail prediction of Housing Prices in Georgia Census Tracts in order to predict Gentrification. More specifically, models will trained on census data from 2010-2014 and then used to predict the 2014-2018 percent change in median house value

The data set can be found here http://www.weaveatlanta.org/weave.html?file=/atlantaneighborhoods.weave. 5 year aggregated data was downloaded from years 2010-2018. Please see Appendix A for a example of a single years csv 

# Required imports
```{r}
library(purrr)
library(readr)
library(dplyr)
library(stats)
library(DataCombine)
library(stringr)
library(MASS)
library(glmnet)
rmse <- function(x,y) sqrt(mean((x-y)^2))
```

# Directory and data attachments
```{r}
files <- list.files(path ='/Users/ved/Desktop/ACS_5year_folder',full.names=TRUE)

data_2010 = read.csv(files[1],skip=1,header=TRUE)
for (i in c(2,19,20)){ data_2010=full_join(data_2010,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2010)[3:length(names(data_2010))] <- paste0(names(data_2010)[3:length(names(data_2010))], ".2010")
# attach(data_2010)

data_2011 = read.csv(files[3],skip=1,header=TRUE)
for (i in c(4,21,22)){ data_2011=full_join(data_2011,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2011)[3:length(names(data_2011))] <- paste0(names(data_2011)[3:length(names(data_2011))], ".2011")
# attach(data_2011)

data_2012 = read.csv(files[5],skip=1,header=TRUE)
for (i in c(6,23,24)){ data_2012=full_join(data_2012,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2012)[3:length(names(data_2012))] <- paste0(names(data_2012)[3:length(names(data_2012))], ".2012")
# attach(data_2012)

data_2013 = read.csv(files[7],skip=1,header=TRUE)
for (i in c(8,25,26)){ data_2013=full_join(data_2013,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2013)[3:length(names(data_2013))] <- paste0(names(data_2013)[3:length(names(data_2013))], ".2013")
# attach(data_2013)

data_2014 = read.csv(files[9],skip=1,header=TRUE)
for (i in c(10,27,28)){ data_2014=full_join(data_2014,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2014)[3:length(names(data_2014))] <- paste0(names(data_2014)[3:length(names(data_2014))], ".2014")
# attach(data_2014)

data_2015 = read.csv(files[11],skip=1,header=TRUE)
for (i in c(12,29,30)){ data_2015=full_join(data_2015,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2015)[3:length(names(data_2015))] <- paste0(names(data_2015)[3:length(names(data_2015))], ".2015")
# attach(data_2015)

data_2016 = read.csv(files[13],skip=1,header=TRUE)
for (i in c(14,31,32)){ data_2016=full_join(data_2016,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2016)[3:length(names(data_2016))] <- paste0(names(data_2016)[3:length(names(data_2016))], ".2016")
# attach(data_2016)

data_2017 = read.csv(files[15],skip=1,header=TRUE)
for (i in c(16,33,34)){ data_2017=full_join(data_2017,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2017)[3:length(names(data_2017))] <- paste0(names(data_2017)[3:length(names(data_2017))], ".2017")
# attach(data_2017)

data_2018 = read.csv(files[17],skip=1,header=TRUE)
for (i in c(18,35,36)){ data_2018=full_join(data_2018,read.csv(files[i],skip=1, header=TRUE), by= "id")}
names(data_2018)[3:length(names(data_2018))] <- paste0(names(data_2018)[3:length(names(data_2018))], ".2018")
# attach(data_2018)

all_data <- Reduce(function(x, y) merge(x, y, all=TRUE), list(data_2010, data_2011, data_2012,data_2013,data_2014,data_2015,data_2016,data_2017, data_2018))
```

# Data cleaning
```{r}
all_data[is.na(all_data)] = 0 #Set nas to 0
i = c(3:13877) # all non string columns (1 and 2 are census tract and name respectively)
all_data[,i] <- apply(all_data[ , i], 2, function(x) as.numeric(as.character(x))) 
all_data[is.na(all_data)] = 0 

names = names(all_data)
bad_names = names[grepl("Margin.of.Error",names, fixed=TRUE)]
ix <- which(names %in% bad_names)
all_data <- all_data[,-ix]
all_data = all_data[, colSums(all_data != 0) > 0]
for (i in (names(all_data)))
  {str_replace(i, '-', '')
    str_replace(i, '\\*', '')
    str_replace(i, '\\+', '')
    }

i = c(3:6719)
all_data[,i] <- apply(all_data[ , i], 2, function(x) as.numeric(as.character(x)))
```

# Individual identification of columns
```{r}
names = names(all_data)
good_names1 = names[grepl('Percent..HOUSING.OCCUPANCY..Vacant.housing.units',names,fixed=TRUE)] #2010-2012
good_names2 = names[grepl('Estimate..HOUSING.OCCUPANCY..Homeowner.vacancy.rate',names,fixed=TRUE)] #2010-2016
good_names3 = names[grepl('Estimate..HOUSING.OCCUPANCY..Total.housing.units.201',names,fixed=TRUE)] #2010-2018 (use only estimate. not percent.estimate)
good_names4 = names[grepl('Percent..HOUSING.OCCUPANCY..Occupied.housing.units',names,fixed=TRUE)] #2010-2012
good_names5 = names[grepl('Percent..HOUSING.TENURE..Renter.occupied',names,fixed=TRUE)] #2010-2012
good_names6 = names[grepl('Percent..VEHICLES.AVAILABLE..No.vehicles.available',names,fixed=TRUE)] #2010-2012
good_names7 = names[grepl('Percent..SELECTED.CHARACTERISTICS..No.telephone.service.available',names,fixed=TRUE)] #2010-2012
good_names8 = names[grepl('Percent..SEX.AND.AGE..Male',names,fixed=TRUE)] #2010-2012
good_names9 = names[grepl('Percent..UNITS.IN.STRUCTURE..1.unit..detached',names,fixed=TRUE)]
good_names10 = names[grepl('Estimate..GROSS.RENT..No.rent.paid',names,fixed=TRUE)]
good_names11 = names[grepl('Percent..YEAR.STRUCTURE.BUILT..Built.2005.or.later',names,fixed=TRUE)]
good_names12 = names[grepl('Estimate..ROOMS..Median.rooms',names,fixed=TRUE)]
good_names13 = names[grepl('Estimate..HOUSING.TENURE..Average.household.size.of.owner.occupied.unit',names,fixed=TRUE)]
good_names14 = names[grepl('Percent..YEAR.HOUSEHOLDER.MOVED.INTO.UNIT..Moved.in.2005.or.later',names,fixed=TRUE)]#14 covers 2010
good_names15 = names[grepl('Percent..HOUSE.HEATING.FUEL..Utility.gas',names,fixed=TRUE)] #16 covers 2010
good_names16 = names[grepl('Percent..SELECTED.CHARACTERISTICS..Lacking.complete.plumbing.facilities',names,fixed=TRUE)] #16 covers 2010
good_names17 = names[grepl('Renter.occupied.housing.units..Estimate..HOUSEHOLD.SIZE..4.or.more.person.household',names,fixed=TRUE)] #17 covers 2010,2014
good_names18 = names[grepl('Renter.occupied.housing.units..Estimate..FAMILY.TYPE.AND.PRESENCE.OF.OWN.CHILDREN..No.related.children.under.18.years',names,fixed=TRUE)] #18 covers 2010,2014
good_names19 = names[grepl('Estimate..VALUE..Median..dollars',names,fixed=TRUE)] #19 covers 2010
good_names20 = names[grepl('Estimate..VALUE..Owner.occupied.units..Median..dollars',names,fixed=TRUE)] #20 covers 2014,2018
good_names21 = names[grepl('Estimate..GROSS.RENT..Median..dollars',names,fixed=TRUE)] #21 covers 2010
good_names22 = names[grepl('Occupied.housing.units..Estimate..HOUSEHOLD.INCOME.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS...Median.household.income..dollars',names,fixed=TRUE)] #22 covers 2014
good_names25 = names[grepl('Percent..YEAR.HOUSEHOLDER.MOVED.INTO.UNIT..Occupied.housing.units..Moved.in.2010.or.later',names,fixed=TRUE)] #25 covers 2014
good_names28 = names[grepl('Percent..HOUSE.HEATING.FUEL..Occupied.housing.units..Utility.gas',names,fixed=TRUE)] #28 covers 2014
good_names29 = names[grepl('Percent..SELECTED.CHARACTERISTICS..Occupied.housing.units..Lacking.complete.plumbing.facilities',names,fixed=TRUE)] #29 covers 2014
good_names33 = names[grepl('Estimate..GROSS.RENT..Occupied.units.paying.rent..Median..dollars',names,fixed=TRUE)] #33 covers 2014,2018
good_names34= names[grepl('Occupied.housing.units..Estimate..HOUSEHOLD.INCOME.IN.THE.PAST.12.MONTHS..IN.2010.INFLATION.ADJUSTED.DOLLARS...Median.household.income..dollars',names,fixed=TRUE)] #34 covers 2010
good_names36 = names[grepl('Percent..HOUSING.OCCUPANCY..Total.housing.units..Vacant.housing.units',names,fixed=TRUE)] #2013-2016
good_names37 = names[grepl('Percent.Estimate..HOUSING.OCCUPANCY..Total.housing.units..Vacant.housing.units',names,fixed=TRUE)] #2017, 2018
good_names38 = names[grepl('Estimate..HOUSING.OCCUPANCY..Total.housing.units..Homeowner.vacancy.rate',names,fixed=TRUE)] #2017,2018 (Use Estimate not Percent.Estimate)
good_names39 = names[grepl('Percent..HOUSING.OCCUPANCY..Total.housing.units..Occupied.housing.units',names,fixed=TRUE)] #2013-2016
good_names40 = names[grepl('Percent.Estimate..HOUSING.OCCUPANCY..Total.housing.units..Occupied.housing.units',names,fixed=TRUE)] #2017-2018
good_names41 = names[grepl('Percent..HOUSING.TENURE..Occupied.housing.units..Renter.occupied',names,fixed=TRUE)] #2013-2016
good_names42 = names[grepl('Percent.Estimate..HOUSING.TENURE..Occupied.housing.units..Renter.occupied.2018',names,fixed=TRUE)] #2018
good_names43 = names[grepl('Percent..VEHICLES.AVAILABLE..Occupied.housing.units..No.vehicles.available',names,fixed=TRUE)] #2013-2016
good_names44 = names[grepl('Percent.Estimate..VEHICLES.AVAILABLE..Occupied.housing.units..No.vehicles.available',names,fixed=TRUE)] #2017-2018
good_names45 = names[grepl('Percent..SELECTED.CHARACTERISTICS..Occupied.housing.units..No.telephone.service.available',names,fixed=TRUE)] #2013-2016
good_names46 = names[grepl('Percent.Estimate..SELECTED.CHARACTERISTICS..Occupied.housing.units..No.telephone.service.available',names,fixed=TRUE)] #2017-2018
good_names47 = names[grepl('Percent..RACE..White',names,fixed=TRUE)] #2010-2012
good_names48 = names[grepl("Percent..RACE..Race.alone.or.in.combination.with.one.or.more.other.races..Total.population..White",names,fixed=TRUE)] #2013-2016
good_names49 = names[grepl("Percent.Estimate..Race.alone.or.in.combination.with.one.or.more.other.races..Total.population..White",names,fixed=TRUE)] #2017-2018
good_names50 = names[grepl('Percent..SEX.AND.AGE..Total.population..Male',names,fixed=TRUE)] #2014,2016
good_names51 = names[grepl('Percent.Estimate..SEX.AND.AGE..Total.population..Male',names,fixed=TRUE)] #2017,2018
good_names52 = names[grepl('Estimate..GROSS.RENT..Occupied.units.paying.rent..No.rent.paid.2018', names,fixed=TRUE)] #for goodanames10 2018
good_names53 = names[grepl('Percent.Margin.of.Error..YEAR.STRUCTURE.BUILT..Total.housing.units..Built.2000.to.2009.2014', names,fixed=TRUE)]  #for goodanames11 2014
good_names54 = names[grepl('Percent.Estimate..YEAR.STRUCTURE.BUILT..Total.housing.units..Built.2014.or.later.2018', names,fixed=TRUE)]  #for goodanames11 2018
good_names55 = names[grepl('Estimate..ROOMS..Total.housing.units..Median.rooms.2014', names,fixed=TRUE)]  #for goodanames12 2014
good_names56 = names[grepl('Percent.Estimate..ROOMS..Total.housing.units..Median.rooms.2018', names,fixed=TRUE)]  #for goodanames12 2018
good_names57 = names[grepl('Estimate..HOUSING.TENURE..Occupied.housing.units..Average.household.size.of.owner.occupied.unit.2018', names,fixed=TRUE)]  #for goodanames13 2018

good_names = c(good_names1,good_names2,good_names3,good_names4,good_names5,good_names6,good_names7,good_names8,good_names9,good_names10,
               good_names11,good_names12,good_names13,good_names14, good_names15,good_names16,good_names17,good_names18,good_names19,good_names20,
               good_names21,good_names22, good_names25, good_names28,good_names29,
               good_names33,good_names34, good_names36, good_names37,good_names38, good_names39, good_names40, 
               good_names41,good_names42, good_names43, good_names44,good_names45, good_names46,good_names47, good_names48, good_names49, good_names50, 
               good_names51,good_names52, good_names53,good_names54, good_names55,good_names56,good_names57,
               'id','Geographic.Area.Name.x','Percent_Change_in_Median_Home_Value_2010_to_2014',
               'Percent_Change_in_Median_Home_Value_2014_to_2018','Percent..YEAR.STRUCTURE.BUILT..Total.housing.units..Built.2010.or.later.2014')

ix <- which(names %in% good_names)
all_data <- all_data[,ix]

#write.csv(names(data_2014),"/Users/ved/Desktop/ACS_5year_folder\\col.csv", row.names = FALSE)
```

# Percent Change Calculations
```{r}
all_data$Percent_Change_in_Median_Home_Value_2010_to_2014 <- (all_data$Estimate..VALUE..Owner.occupied.units..Median..dollars..2014 - all_data$Estimate..VALUE..Median..dollars..2010)/(all_data$Estimate..VALUE..Median..dollars..2010)*100
all_data$Percent_Change_in_Median_Home_Value_2014_to_2018 <- (all_data$Estimate..VALUE..Owner.occupied.units..Median..dollars..2018 - all_data$Estimate..VALUE..Owner.occupied.units..Median..dollars..2014)/(all_data$Estimate..VALUE..Owner.occupied.units..Median..dollars..2014)*100
all_data$Percent_Change_in_No_Rent_Paid <- (all_data$Estimate..GROSS.RENT..No.rent.paid.2014 - all_data$Estimate..GROSS.RENT..No.rent.paid.2010)
all_data$Percent_Change_in_Built_2009_2005 <-(all_data$Percent..YEAR.STRUCTURE.BUILT..Total.housing.units..Built.2010.or.later - all_data$Percent..YEAR.STRUCTURE.BUILT..Built.2005.or.later.2010)
all_data$Percent_Change_in_Median_Rooms <- 100*((all_data$Estimate..ROOMS..Total.housing.units..Median.rooms.2014 - all_data$Estimate..ROOMS..Median.rooms.2010)/all_data$Estimate..ROOMS..Median.rooms.2010)
all_data$Percent_Change_in_Average_Household_Size_Owner <- 100*((all_data$Estimate..HOUSING.TENURE..Average.household.size.of.owner.occupied.unit.2014 - all_data$Estimate..HOUSING.TENURE..Average.household.size.of.owner.occupied.unit.2010)/all_data$Estimate..HOUSING.TENURE..Average.household.size.of.owner.occupied.unit.2010)
all_data$Percent_Change_in_Recent_Move_In_2010_to_2014 <- (all_data$Percent..YEAR.HOUSEHOLDER.MOVED.INTO.UNIT..Occupied.housing.units..Moved.in.2010.or.later.2014- all_data$Percent..YEAR.HOUSEHOLDER.MOVED.INTO.UNIT..Moved.in.2005.or.later.2010)/(all_data$Percent..YEAR.HOUSEHOLDER.MOVED.INTO.UNIT..Moved.in.2005.or.later.2010)*100
all_data$Percent_Change_in_Gas_Usage_2010_to_2014 <- (all_data$Percent..HOUSE.HEATING.FUEL..Occupied.housing.units..Utility.gas.2014-all_data$Percent..HOUSE.HEATING.FUEL..Utility.gas.2010)/(all_data$Percent..HOUSE.HEATING.FUEL..Utility.gas.2010)*100
all_data$Percent_Change_in_Plumbing_2010_to_2014 <- (all_data$Percent..SELECTED.CHARACTERISTICS..Occupied.housing.units..Lacking.complete.plumbing.facilities.2014-all_data$Percent..SELECTED.CHARACTERISTICS..Lacking.complete.plumbing.facilities.2010)
all_data$Percent_Change_in_Large_Households_2010_to_2014 <- (all_data$Renter.occupied.housing.units..Estimate..HOUSEHOLD.SIZE..4.or.more.person.household.2014 - all_data$Renter.occupied.housing.units..Estimate..HOUSEHOLD.SIZE..4.or.more.person.household.2010)/(all_data$Renter.occupied.housing.units..Estimate..HOUSEHOLD.SIZE..4.or.more.person.household.2010)*100
all_data$Percent_Change_in_Children_2010_to_2014 <- (all_data$Renter.occupied.housing.units..Estimate..FAMILY.TYPE.AND.PRESENCE.OF.OWN.CHILDREN..No.related.children.under.18.years.2014 - all_data$Renter.occupied.housing.units..Estimate..FAMILY.TYPE.AND.PRESENCE.OF.OWN.CHILDREN..No.related.children.under.18.years.2010)/(all_data$Renter.occupied.housing.units..Estimate..FAMILY.TYPE.AND.PRESENCE.OF.OWN.CHILDREN..No.related.children.under.18.years.2010)*100
all_data$Percent_Change_in_Rent_2010_to_2014 <- (all_data$Estimate..GROSS.RENT..Occupied.units.paying.rent..Median..dollars..2014-all_data$Estimate..GROSS.RENT..Median..dollars..2010)/(all_data$Estimate..GROSS.RENT..Median..dollars..2010)*100
all_data$Percent_Change_in_Income_2010_to_2014 <-(all_data$Occupied.housing.units..Estimate..HOUSEHOLD.INCOME.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS...Median.household.income..dollars..2014-all_data$Occupied.housing.units..Estimate..HOUSEHOLD.INCOME.IN.THE.PAST.12.MONTHS..IN.2010.INFLATION.ADJUSTED.DOLLARS...Median.household.income..dollars..2010)/ (all_data$Occupied.housing.units..Estimate..HOUSEHOLD.INCOME.IN.THE.PAST.12.MONTHS..IN.2010.INFLATION.ADJUSTED.DOLLARS...Median.household.income..dollars..2010)*100
all_data$Percent_Change_in_Vacant_Housing_Units_2010_to_2014 <- (all_data$Percent..HOUSING.OCCUPANCY..Total.housing.units..Vacant.housing.units.2014 - all_data$Percent..HOUSING.OCCUPANCY..Vacant.housing.units.2010)/all_data$Percent..HOUSING.OCCUPANCY..Vacant.housing.units.2010 * 100
all_data$Percent_Change_in_Occupied_Housing_Units_2010_to_2014 <- (all_data$Percent..HOUSING.OCCUPANCY..Total.housing.units..Occupied.housing.units.2014 - all_data$Percent..HOUSING.OCCUPANCY..Occupied.housing.units.2010) / all_data$Percent..HOUSING.OCCUPANCY..Occupied.housing.units.2010 * 100
all_data$Percent_Change_in_Total_Housing_Units_2010_to_2014 <- (all_data$Estimate..HOUSING.OCCUPANCY..Total.housing.units.2014 - all_data$Estimate..HOUSING.OCCUPANCY..Total.housing.units.2010) / all_data$Estimate..HOUSING.OCCUPANCY..Total.housing.units.2010 * 100
all_data$Percent_Change_in_Homeowner_Vacancy_Rate_2010_to_2014 <- (all_data$Estimate..HOUSING.OCCUPANCY..Homeowner.vacancy.rate.2014 - all_data$Estimate..HOUSING.OCCUPANCY..Homeowner.vacancy.rate.2010) / all_data$Estimate..HOUSING.OCCUPANCY..Homeowner.vacancy.rate.2010 * 100
all_data$Percent_Change_in_Renter_Occupied_Housing_Tenure_2010_to_2014 <- (all_data$Percent..HOUSING.TENURE..Occupied.housing.units..Renter.occupied.2014 - all_data$Percent..HOUSING.TENURE..Renter.occupied.2010) / all_data$Percent..HOUSING.TENURE..Renter.occupied.2010 * 100
all_data$Percent_Change_in_No_Vehicles_Available_2010_to_2014 <- (all_data$Percent..VEHICLES.AVAILABLE..Occupied.housing.units..No.vehicles.available.2014- all_data$Percent..VEHICLES.AVAILABLE..No.vehicles.available.2010)/all_data$Percent..VEHICLES.AVAILABLE..No.vehicles.available.2010 * 100
all_data$Percent_Change_in_No_Telephone_Service_Availble_2010_to_2014 <- (all_data$Percent..SELECTED.CHARACTERISTICS..Occupied.housing.units..No.telephone.service.available.2014- all_data$Percent..SELECTED.CHARACTERISTICS..No.telephone.service.available.2010) / all_data$Percent..SELECTED.CHARACTERISTICS..No.telephone.service.available.2010 * 100
all_data$Percent_Change_in_Males_2010_to_2014 <- (all_data$Percent..SEX.AND.AGE..Total.population..Male.2014- all_data$Percent..SEX.AND.AGE..Male.2010)/all_data$Percent..SEX.AND.AGE..Male.2010*100
all_data$Percent_Change_in_Whites_2010_to_2014 <- (all_data$Percent..RACE..Race.alone.or.in.combination.with.one.or.more.other.races..Total.population..White.2014- all_data$Percent..RACE..White.2010)/all_data$Percent..RACE..White.2010*100
```

# Individual column NA adjustments
```{r}
all_data <- all_data[!is.na(all_data$Percent_Change_in_Median_Home_Value_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Median_Home_Value_2014_to_2018),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Recent_Move_In_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Plumbing_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Large_Households_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Children_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Rent_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Income_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Vacant_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Occupied_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Total_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Homeowner_Vacancy_Rate_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Renter_Occupied_Housing_Tenure_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_No_Vehicles_Available_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_No_Telephone_Service_Availble_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Males_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Whites_2010_to_2014),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_No_Rent_Paid),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Median_Rooms),]
all_data <- all_data[!is.na(all_data$Percent_Change_in_Average_Household_Size_Owner),]
```

# Individual column infinity adjustments
```{r}
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Median_Home_Value_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Median_Home_Value_2014_to_2018),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Recent_Move_In_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Plumbing_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Large_Households_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Children_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Rent_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Income_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Vacant_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Occupied_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Total_Housing_Units_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Homeowner_Vacancy_Rate_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Renter_Occupied_Housing_Tenure_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_No_Vehicles_Available_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_No_Telephone_Service_Availble_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Males_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Whites_2010_to_2014),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_No_Rent_Paid),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Median_Rooms),]
all_data <- all_data[!is.infinite(all_data$Percent_Change_in_Average_Household_Size_Owner),]
```

Given a vast amount of data, we will identify significant variables using an X regression
# Definining predicted data set
```{r}
names = names(all_data)
pred_names <- c(names[grepl('Percent_Change_in',names,fixed=TRUE)],'id','Geographic.Area.Name.x')
ix <- which(names %in% pred_names)
pred_data <- all_data[,ix]
```

# Shape/ Summary of Data
```{r}

```

# Initial multiple linear regression
```{r}
full.model <- lm(pred_data$Percent_Change_in_Median_Home_Value_2014_to_2018~., na.action=na.exclude, data=pred_data[,c(-1,-2,-4)],x=TRUE)
rmse(predict(full.model,pred_data[,-4]), pred_data$Percent_Change_in_Median_Home_Value_2014_to_2018)
plot(resid(full.model))
sum(coef(full.model)!=0)
sum(abs(coef(full.model)))
avplots(full.model)
```

# Stepwise Regression
```{r}
g.step <- step(full.model)
summary(g.step)
rmse(predict(g.step,pred_data), pred_data$Percent_Change_in_Median_Home_Value_2014_to_2018)

```
As seen above x,y,z are significant. Carried forward with interactions in stepwise regression

# Stepwise Regression 
```{r}
g_stepmodel = lm(pred_data$Percent_Change_in_Median_Home_Value_2014_to_2018~
              (pred_data$Percent_Change_in_No_Rent_Paid+
              pred_data$Percent_Change_in_Built_2009_2005+
              pred_data$Percent_Change_in_Recent_Move_In_2010_to_2014+
              pred_data$Percent_Change_in_Plumbing_2010_to_2014+
              pred_data$Percent_Change_in_Large_Households_2010_to_2014+
              pred_data$Percent_Change_in_Children_2010_to_2014+
              pred_data$Percent_Change_in_Occupied_Housing_Units_2010_to_2014+
              pred_data$Percent_Change_in_Total_Housing_Units_2010_to_2014+
              pred_data$Percent_Change_in_Homeowner_Vacancy_Rate_2010_to_2014+
              pred_data$Percent_Change_in_Males_2010_to_2014+
              pred_data$Percent_Change_in_Whites_2010_to_2014+
              pred_data$Percent_Change_in_Income_2010_to_2014)^2 ,na.action=na.exclude, data=pred_data[,c(-1,-2,-4)],x=TRUE)
g_stepinteractionmodel <- step(g_stepmodel)
summary(g_stepinteractionmodel)
rmse(predict(g_stepinteractionmodel,pred_data), pred_data$Percent_Change_in_Median_Home_Value_2014_to_2018)

```

# Conclusion
Based on our final stepwise model with interactions, we find that 80.79% of the variability in housing prices can be explained by changes in people not paying rent, people moving in, access to plumbing, number of large households, percent of families with children, number of occupied housing units, number of total housing units, homeowner vacancy rate, percentage of males, percentage of whites, and income, as well as interactions between these variables.

# Appendix
## A

